Measuring precise radial velocities and cross-correlation function line-profile variations using a Skew Normal density\thanks{Based on observations collected at the La Silla Parana Observatory, ESO (Chile), with the HARPS spectrograph at the 3.6-m telescope.}

### 2018-07-06 manuscript.tex
edited sec 4.3

\begin{figure*}[htbp]
\begin{center}
\includegraphics[height = 6in]{Spot_NEW_CORRECTION_[3]CorrectionActivity_RadialVelocity_vs_time.pdf} 
   \caption{(top) The spurious estimated RVs (black dots) caused by a spot in the simulated data, the estimated RVs using Eq.~\ref{eq:RV:correction} (red crosses) and the estimated RVs using the usual correction for stellar activity (green triangles), based on $RV_{\text{activity}}=\beta_0+\beta_1 \gamma + \beta_2 \text{SN FWHM}$ for the SN fit and on $RV_{\text{activity}}=\beta_0+\beta_1 \text{BIS SPAN} + \beta_2 \text{FWHM}$ for the normal fit.
   %
 (bottom) The residuals from the model fit using Eq.~\ref{eq:RV:correction} (New corr. std, red crosses) and the residuals from the usual correction (Usual corr. std, green triangles). The residuals have a smaller systematic component when using the proposed model of 
  Eq.~\ref{eq:RV:correction} (red crosses) compared to the usual model (green triangles).
The tests of statistical significance on the parameters are presented in Table \ref{table:spot.test}.
}
    \label{fig:spot.correction}
\end{center}
\end{figure*}

@@

The final simulation presented here includes a planetary signal influencing the CCF along with the 1\% spot modelled previously (see Sec.~\ref{sec:soap.spot}). 
The purpose of this example is to check if we are able to disentangle these two different sources of variations when using the parameters derived using a Normal or a SN fit to the CCF. In this scenario the planet is injected with a semi-amplitude of 10 \ms with no eccentricity and with a period corresponding to one third of the stellar rotational period, i.e. one-third of 25 days.

@@

The correlation between the different CCF parameters are displayed in Fig.~\ref{fig:spot.plus.planet.corr}. 
The correlations are weaker than in the case of the spot due to the planet inducing changes in RV without affecting the width or the asymmetry of the CCF.  However, the strength of the correlations between the CCF asymmetry parameters and RV are comparable with the ones obtained for the spot-only model: $\gamma$--SN median RV has the strongest correlation ($R=-0.84$), followed by the correlation between BIS SPAN--N mean RV (($R=-0.78$)) and finally by the correlation between $\gamma$--SN mean RV ($R=-0.75$). 
The patterns observed in the width-RV phase space correlations in Fig.~\ref{fig:spot.plus.planet.corr} follow a circle, similar to the spot-only model; no correlation is observed between those two parameters.

\begin{figure}[htbp]
\begin{center}
\includegraphics[width = 3.6in]{RV_comparison_SPOT_PLANET.pdf} 
%\includegraphics[height = 2.5in]{RV_se_comparison_SPOT.pdf} 
 \caption{RV estimates for N mean RV (red dashed line),  SN mean RV (black line) or SN median RV (cyan dots-dashed line). In this case, the CCFs have been generated using SOAP 2.0, considering an equatorial 1\% spot on the simulated Sun in addition to a planet with a period of one third of the rotational period of the star and with an amplitude of 10\,\ms. The star do one full rotation between phase -0.5 and 0.5, with the spot being seen face-on for phase 0.}
   %\caption{(left) RVs changes as function of the orbital phase in the case in which a spot is present on the photosphere of the star and a planet is injected. N mean RV seems to have the largest variations caused by the combined action of spot and planet. (right) Evaluation of the standard errors corresponding to the defined RVs. The standard errors retrieved for SN median RV are $10 \%$ smaller than the standard errors derived for RV. SN mean RV has the largest related uncertainties.}
    \label{fig:spot.plus.planet}
\end{center}
\end{figure}

\begin{figure*}[htbp]
\begin{center}
\includegraphics[height = 6in]{SOAP_SPOT_PLANET_Comparison_para_SN.pdf} 
   \caption{Evaluation of the correlation between the RVs and the asymmetry parameters of the simulated data with a 1\% spot and an injected planetary signal.  The shape of the CCF changes as the spot moves, producing statistically significant correlations only between the estimated RVs and the asymmetry parameter. The correlations between the estimated RVs and the width parameter of the CCF is weaker than the case with only a spot.}
    \label{fig:spot.plus.planet.corr}
\end{center}
\end{figure*}

In order to correct the estimated RVs from the spurious variation caused by the spot, the proposed model for correcting the activity is added to a signal that takes into account the RV variations caused by a injected planet. The observed RVs can therefore be modelled as a combination of the activity and the planetary signals:
%
\begin{equation}
RV= RV_{\text{activity}} + RV_{\text{planet}},
\label{eq:RV:correction.overall}
\end{equation}
%
where $RV_{\text{activity}}$ can be found in Eq.~\ref{eq:RV:correction}, and $RV_{\text{planet}}$, in the case with no eccentricity, can be modelled by the following sinusoidal function:
%
\begin{equation}
RV_{\text{exoplanet}}= K \sin \left(\frac{2 \pi}{P} (t - t_{0})\right),
\label{eq:RV:correction.planet}
\end{equation}
%
with amplitude $K$, orbital period $P$, and an epoch at the periapsis $t_{0}$.  The previous three unknown parameters define the planetary orbit.

%We note that the p--value associated with the amplitude parameter $K$ is particularly relevant for rejecting or not rejecting the assumption about the presence of an orbiting companion. Moreover we note also that Eq.~\ref{eq:RV:correction.planet} is highly non linear, meaning that the estimation of all the parameters involved in Eq.~\ref{eq:RV:correction.overall} has to be done numerically. %We used non linear least squares and the results are displayed in Fig.~\ref{fig:spotplanet.correction}. We can see how in this case we are able to disentangle the spurious variations in RVs caused by stellar activity from the pure Doppler-shift due to the planet. 
The proposed model from Eq.~\ref{eq:RV:correction.overall} was fitted to the RV data and the results of the estimated model are summarized in Table \ref{table:spotplanet.test}.   Except for the width parameters with coefficient $\beta_3$, all the other Normal or SN parameters are significantly useful to explain the RV variation induced by a spot plus a planet. We also observe that the RV residuals, once corrected for stellar activity and the presence of the planet, are comparable in terms of rms for all the three different RV estimates, with SN mean RV giving a slightly smaller value.
%\begin{figure*}[htbp]
%   \centering
%\includegraphics[height = 4in]{SPOT_PLANET_NEW_CORRECTION.pdf} 
%   \caption{Set of  variations in RVs estimated using a Normal and a SN fit before and once corrected from stellar activity. In this case there are spurious variations caused by the spot and pure Doppler-shift due to the planet. The correction is done using Eq.~\ref{eq:RV:correction.overall} and the estimated parameters are presented in Table \ref{table:spotplanet.test}. In this case, by solving Eq.~\ref{eq:RV:correction.overall}, we are able to completely disentangle the spurious variations in RVs caused by the presence of the spot from the pure dopplershift caused by the exoplanet.}
%    \label{fig:spotplanet.correction}
%\end{figure*}

\begin{table}
\begin{center}
\caption{P--values for the different coefficients used in Eq.~\ref{eq:RV:correction} for the correction of stellar activity induced by an equatorial 1\% spot on the simulated Sun, and a planet with period one third of the rotational period and a semi-amplitude 10\,\ms. All the parameters corresponding to the Normal or SN variables are statistically significant to explain the spurious RV variations caused by this spot plus planet, except for the width of the CCF. Note that since nonlinear least squares was required, the residual standard error rather than the $R^2$ is displayed as a reference.}
\label{table:spotplanet.test}
\begin{tabular}{|c|c|c|c|}
\hline
Parameter          & N mean RV         &   SN mean RV &   SN median RV \\
\hline
$\beta_{0}$            &    $0.00063$    & $2e-16$  & $1.42e-09$ \\
\hline
$\beta_{1}$            &    $2e-16$    & $2e-16$  & $2e-16$ \\
\hline
$\beta_{2}$            &     $2e-16$   & $2e-16$ & $2e-16$\\
\hline
$\beta_{3}$            &     $0.067$   &  $0.40$  & $0.38$\\
\hline
$\beta_{4}$            &     $2e-16$   &  $2e-16$ & $2e-16$\\
\hline
K            &     $2e-16$   &  $2e-16$   & $2e-16$ \\
\hline
P            &     $2e-16$   &  $2e-16$ & $2e-16$ \\
\hline
$t_{0}$            &     $2e-16$   &  $2e-16$ & $2e-16$ \\
\hline
$\text{Residuals}$      &     $0.71 \ms$    &  $ 0.66 \ms$ & $0.70 \ms$  \\
\hline
\end{tabular}
\end{center}
\end{table}

>>>>

\begin{figure*}[htbp]
\begin{center}
\includegraphics[height = 6in]{Spot_NEW_CORRECTION_[3]CorrectionActivity_RadialVelocity_vs_time.pdf} 
   \caption{(top) The spurious estimated RVs (black dots) caused by a spot in the simulated data, the estimated RVs using Eq.~\ref{eq:RV:correction} (red crosses) and the estimated RVs using the usual correction for stellar activity (green triangles), based on $RV_{\text{activity}}=\beta_0+\beta_1 \gamma + \beta_2 \text{SN FWHM}$ for the SN fit and on $RV_{\text{activity}}=\beta_0+\beta_1 \text{BIS SPAN} + \beta_2 \text{FWHM}$ for the normal fit.
   %
 (bottom) The residuals from the model fit using Eq.~\ref{eq:RV:correction} (red crosses) and the residuals from the usual correction (green triangles). 
 The std. are also reported in the legend, and the residuals have a smaller systematic component when using the proposed model compared to the usual model.
The tests of statistical significance on the parameters are presented in Table~\ref{table:spot.test}.
}
    \label{fig:spot.correction}
\end{center}
\end{figure*}

@@

The final simulation includes a planetary signal influencing the CCF along with the 1\% spot modelled previously (see Sec.~\ref{sec:soap.spot}). 
The purpose of this example is to check if we are able to disentangle these two different sources of variations when using the parameters derived using a Normal or a SN model for the CCF. In this scenario the planet is injected with a semi-amplitude of 10 \ms with no eccentricity and with a period corresponding to one-third of the stellar rotational period, i.e. one-third of 25 days.

@@

The correlation between the different CCF parameters are displayed in Fig.~\ref{fig:spot.plus.planet.corr}. 
The correlations are weaker than in the case of the spot due to the planet inducing changes in RV without affecting the width or the asymmetry of the CCF.  However, the order of the strength of the correlations between the CCF asymmetry parameters and RV are comparable with the ones obtained for the spot-only model: $\gamma$--SN median RV has the strongest correlation ($R=-0.84$), followed by the correlation between BIS SPAN--N mean RV ($R=-0.78$) and finally by the correlation between $\gamma$--SN mean RV ($R=-0.75$). 
The patterns observed in the width-RV phase space correlations in Fig.~\ref{fig:spot.plus.planet.corr} follow a circle, similar to the spot-only model; no substantial correlation is observed between those two parameters.

\begin{figure}[htbp]
\begin{center}
\includegraphics[width = 3.6in]{RV_comparison_SPOT_PLANET.pdf} 
%\includegraphics[height = 2.5in]{RV_se_comparison_SPOT.pdf} 
 \caption{RV estimates for N mean RV (red dashed line),  SN mean RV (black line) or SN median RV (cyan dotted-dashed line). In this case, the CCFs have been generated using SOAP 2.0, considering an equatorial 1\% spot on the simulated Sun in addition to a planet with a period of one-third of the rotational period of the star and with an amplitude of 10\,\ms. The star does one full rotation between phase -0.5 and 0.5, with the spot being seen face-on at phase 0.}
   %\caption{(left) RVs changes as function of the orbital phase in the case in which a spot is present on the photosphere of the star and a planet is injected. N mean RV seems to have the largest variations caused by the combined action of spot and planet. (right) Evaluation of the standard errors corresponding to the defined RVs. The standard errors retrieved for SN median RV are $10 \%$ smaller than the standard errors derived for RV. SN mean RV has the largest related uncertainties.}
    \label{fig:spot.plus.planet}
\end{center}
\end{figure}

\begin{figure*}[htbp]
\begin{center}
\includegraphics[height = 6in]{SOAP_SPOT_PLANET_Comparison_para_SN.pdf} 
   \caption{Evaluation of the correlation between the RVs and the asymmetry parameters of the simulated data with a 1\% spot and an injected planetary signal.  The shape of the CCF changes as the spot moves, producing statistically significant correlations only between the estimated RVs and the asymmetry parameter. The correlations between the estimated RVs and the width parameter of the CCF are weaker than the case with only a spot.}
    \label{fig:spot.plus.planet.corr}
\end{center}
\end{figure*}

In order to correct the estimated RVs from the spurious variation caused by the spot, the proposed model for correcting the activity is added to a planetary signal model that takes into account the RV variations caused by a planet. The observed RVs can therefore be modelled as a combination of the activity and the planetary signals:
%
\begin{equation}
RV= RV_{\text{activity}} + RV_{\text{planet}},
\label{eq:RV:correction.overall}
\end{equation}
%
where $RV_{\text{activity}}$ can be found in Eq.~\ref{eq:RV:correction}, and $RV_{\text{planet}}$, in the case with no eccentricity, can be modelled by the following sinusoidal function:
%
\begin{equation}
RV_{\text{exoplanet}}= K \sin \left(\frac{2 \pi}{P} (t - t_{0})\right),
\label{eq:RV:correction.planet}
\end{equation}
%
with amplitude $K$, orbital period $P$, and an epoch at the periapsis $t_{0}$.  The previous three unknown parameters define the planetary orbit.

%We note that the p--value associated with the amplitude parameter $K$ is particularly relevant for rejecting or not rejecting the assumption about the presence of an orbiting companion. Moreover we note also that Eq.~\ref{eq:RV:correction.planet} is highly non linear, meaning that the estimation of all the parameters involved in Eq.~\ref{eq:RV:correction.overall} has to be done numerically. %We used non linear least squares and the results are displayed in Fig.~\ref{fig:spotplanet.correction}. We can see how in this case we are able to disentangle the spurious variations in RVs caused by stellar activity from the pure Doppler-shift due to the planet. 
The proposed model from Eq.~\ref{eq:RV:correction.overall} was fitted to the RV data and the results of the estimated model are summarized in Table~\ref{table:spotplanet.test}.   Except for the width parameters with coefficient $\beta_3$, all the other Normal or SN parameters are significantly useful to explain the RV variation induced by a spot plus a planet. We also observe that the RV residuals, once corrected for stellar activity and the presence of the planet, are comparable in terms of rms for all the three different RV estimates, with SN mean RV giving a slightly smaller value.
%\begin{figure*}[htbp]
%   \centering
%\includegraphics[height = 4in]{SPOT_PLANET_NEW_CORRECTION.pdf} 
%   \caption{Set of  variations in RVs estimated using a Normal and a SN fit before and once corrected from stellar activity. In this case there are spurious variations caused by the spot and pure Doppler-shift due to the planet. The correction is done using Eq.~\ref{eq:RV:correction.overall} and the estimated parameters are presented in Table \ref{table:spotplanet.test}. In this case, by solving Eq.~\ref{eq:RV:correction.overall}, we are able to completely disentangle the spurious variations in RVs caused by the presence of the spot from the pure dopplershift caused by the exoplanet.}
%    \label{fig:spotplanet.correction}
%\end{figure*}

\begin{table}
\begin{center}
\caption{P-values for the different coefficients used in Eq.~\ref{eq:RV:correction} for the correction of stellar activity induced by an equatorial 1\% spot on the simulated Sun, and a planet with a period of one-third the rotational period of the star and a semi-amplitude  of 10\,\ms. All the parameters corresponding to the Normal or SN variables are statistically significant, except the width of the CCF. Note that since nonlinear least squares was required, the residual standard error rather than the $R^2$ is displayed as a reference.}
\label{table:spotplanet.test}
\begin{tabular}{|c|c|c|c|}
\hline
Parameter          & N mean RV         &   SN mean RV &   SN median RV \\
\hline
$\beta_{0}$            &    $0.00063$    & $2e-16$  & $1.42e-09$ \\
\hline
$\beta_{1}$            &    $2e-16$    & $2e-16$  & $2e-16$ \\
\hline
$\beta_{2}$            &     $2e-16$   & $2e-16$ & $2e-16$\\
\hline
$\beta_{3}$            &     $0.067$   &  $0.40$  & $0.38$\\
\hline
$\beta_{4}$            &     $2e-16$   &  $2e-16$ & $2e-16$\\
\hline
K            &     $2e-16$   &  $2e-16$   & $2e-16$ \\
\hline
P            &     $2e-16$   &  $2e-16$ & $2e-16$ \\
\hline
$t_{0}$            &     $2e-16$   &  $2e-16$ & $2e-16$ \\
\hline
$\text{Residuals}$      &     $0.71 \ms$    &  $ 0.66 \ms$ & $0.70 \ms$  \\
\hline
\end{tabular}
\end{center}
\end{table}
